%%=============================================================================
%% Voorwoord
%%=============================================================================

\chapter*{Preface}
\label{ch:voorwoord}

%% TODO:
%% Het voorwoord is het enige deel van de bachelorproef waar je vanuit je
%% eigen standpunt (``ik-vorm'') mag schrijven. Je kan hier bv. motiveren
%% waarom jij het onderwerp wil bespreken.
%% Vergeet ook niet te bedanken wie je geholpen/gesteund/... heeft

This bachelor thesis was written as part of the course which shares the same name. This thesis together with the internship I'm currently doing are the final two courses to complete a bachelor diploma in Applied Information Technology.
\section*{Choosing the subject}
The subject of the thesis is one I picked from a list of suggestions that were posted online on the school platform \textit{Chamilo}. The subject caught my eye because of my personal experience with the issue. There would be times where we would take digital examinations for certain courses and people all around me would be cheating to get better results. I always found it rather disrespectful and disturbing that people were doing this. Furthermore that they were able to do this so easily. It's a real demotivation for students who actually study for the examinations. Because when you get your grades and realize that everybody who cheated got better results than you, then why bother studying if no penalties are given out anyway.\\
This of course devalues the examination itself and the course as a whole. Resulting in people not taking courses with digital exams seriously and getting high grades without ever understanding what the course was really about.\\
So when the opportunity was there for me to possibly change this and help the school with solving the problem, I eagerly picked this as my subject for my bachelor thesis.\\
\section*{Acknowledgements}
Mainly I'd like to thank the different communities on the web that helped me in many situations. Communities like \textit{Serverfault}\footnote{https://serverfault.com/} and \textit{reddit}\footnote{https://www.reddit.com/r/networking/}. They provided me with opinions and insights into certain fields where I myself do not have a lot of knowledge in.\\
Next up I'd like to thanks my mother for buying enough coffee supplies and my father for his unrelenting stimuli. My brother as well for his input on the occasional English translation or some Latex related error.\\
And finally I'd like to thank my promotor and co-promotor for aiding me in building this work and making it a better whole.